---
title: "Ex-8-1: Make tables with `flextable`"
format:
  html:
    toc: true
    toc-location: left
    number-sections: true
webr:
  packages: ['dplyr', 'ggplot2', 'datasets', 'flextable']
  cell-options:
    editor-font-scale: 0.7
    out-width: 100%
filters:
  - webr
---

```{r}
#| include: false 
library(dplyr)
library(ggplot2)
library(flextable)
```

## Exercise 1

**Dataset:** `mtcars`

1.  Load the `mtcars` dataset into R.
2.  Use the `flextable` package to create a table that displays the first 10 rows of the dataset for `mpg`, `cyl`, `disp`, and `hp`.
3.  Apply styles to the table to make
    1. the texts in the header bold using `bold()`
    2. the texts of the `cyl` column center-aligned including the `cyl` in the header using `align()`
    3. the texts of the `hp` column colored red using `color()`
    4. the font size of the texts of the `mpg` 18 if the value of `disp` is greater than 200 using `fontsize()`.

Here is the finished table you are trying to make:

```{r}
#| echo: false
(
ft <- 
  flextable(
    mtcars[1:10, ],
    col_keys = c("mpg", "cyl", "disp", "hp")
  ) %>%
  bold(part = "header") %>%
  align(j = 2, align = "center", part = "all") %>%
  color(j = 4, color = "red") %>%
  fontsize(i = ~ disp > 200, j = 1, size = 18)
) 
```



::: {.panel-tabset}
### Work here
```{webr-r}
#| results: asis
data(mtcars)


```

### Answer
```{r, eval = FALSE}
#| code-fold: true
# Create a simple flextable
ft <- 
  flextable(
    mtcars[1:10, ],
    col_keys = c("mpg", "cyl", "disp", "hp")
  ) %>%
  bold(part = "header") %>%
  align(j = 2, align = "center", part = "all") %>%
  color(j = 4, color = "red")
```

:::


## Exercise 2

**Dataset:** `iris`

1.  Load the `iris` dataset.
2.  Create a `flextable` that displays the first 10 rows of the dataset.
3.  Apply conditional formatting: 
    1.  Change the background color of the  cells in the `Sepal.Length` column where their values are greater than 5.0 in blue using `color` using `bg()`.
    2.  Draw orange borders at the bottom of  the cells in the `Sepal.Width` column if their values are less than 3.0 using `border()`.
    3.  Draw red borders at both sides of the cells in the `Sepal.Width` column if their corresponding values of `Petal.Width` are less than 1.5 using `border()`.

Here is the finished table you are trying to make:

```{r}
#| echo: false
(
ft <- 
  flextable(iris[1:10,]) %>%
  # Apply conditional formatting to Sepal.Length column
  bg(i = ~ Sepal.Length > 5, j = "Sepal.Length", bg = "blue") %>%
  border(i = ~ Sepal.Width > 3, j = ~ Sepal.Width, border.bottom = officer::fp_border(color = "orange"))
)
```


::: {.panel-tabset}
### Work here
```{webr-r}
# Load iris dataset
data(iris)
```

### Answer
```{r, eval = FALSE}
#| code-fold: true
# Create flextable for first 10 rows
ft <- 
  flextable(iris[1:10,]) %>%
  # Apply conditional formatting to Sepal.Length column
  bg(i = ~ Sepal.Length > 5, j = "Sepal.Length", bg = "blue") %>%
  border(i = ~ Sepal.Width > 3, j = ~ Sepal.Width, border.bottom = officer::fp_border(color = "orange"))
```

:::


## Exercise 3: Merging and Grouping Columns

**Dataset:** `airquality`

1.  Load the `airquality` dataset and filter the data so that you have only the first 10 days of May and June.
2.  Create a flextable using selected columns: `Month`, `Day`, `Ozone`, `Wind`.
3.  Group the table by the `Month` column and merge the cells of the `Month` column using `merge_v()`.
4.  Draw a black border to separate May and June observations
5.  Draw a black line at the bottom of the table which was somehow lost in step 3 using `fix_border_issues()`.
6.  Add a caption to the table using `set_caption()`.

Here is the finished table you are trying to make:

```{r}
#| echo: false
(
ft <- 
  airquality %>%
  dplyr::filter(Month %in% c(5, 6) & Day <= 10) %>%
  .[, c("Month", "Day", "Ozone", "Wind")] %>%
  # Create flextable
  flextable() %>%
  # Merge Month column
  merge_v(j = "Month") %>%
  border(i = 10, border.bottom = officer::fp_border(color = "black")) %>%
  fix_border_issues() %>%
  # Add a caption
  set_caption("Airquality Data Table with Grouped Month")
) 
```

::: {.panel-tabset}
### Work here
```{webr-r}
# Load airquality dataset
data(airquality)
```

### Answer
```{r, eval = FALSE}
#| code-fold: true
# Select columns
ft <- 
  airquality %>%
  dplyr::filter(Month %in% c(5, 6) & Day <= 10) %>%
  .[, c("Month", "Day", "Ozone", "Wind")] %>%
  # Create flextable
  flextable() %>%
  # Merge Month column
  merge_v(j = "Month") %>%
  border(i = 10, border.bottom = officer::fp_border(color = "black")) %>%
  fix_border_issues() %>%
  # Add a caption
  set_caption("Airquality Data Table with Grouped Month")
```

:::

## Exercise 4: Custom Header and Footer

**Dataset:** `iris`

1.  Load the `iris` dataset.
2.  Create a table with the first 10 rows of the dataset.
3.  Modify the header by renaming columns `Sepal.Length` and `Sepal.Width` to `Sepal Length` and `Sepal Width`, respectively.
4.  Add a custom footer explaining what the dataset represents.

Here is the finished table you are trying to make:

```{r}
#| echo: false
(
ft <- 
  flextable(iris[1:10,]) %>%
  # Modify header labels
  set_header_labels(Sepal.Length = "Sepal Length", Sepal.Width = "Sepal Width") %>%
  # Add a footer
  add_footer_lines("Iris dataset: Measurements of iris flowers from three species") 
)
```

::: {.panel-tabset}
### Work here
```{webr-r}
# Load iris dataset
data(iris)
```

### Answer
```{r, eval = FALSE}
#| code-fold: true
ft <- 
  flextable(iris[1:15,]) %>%
  # Modify header labels
  set_header_labels(Sepal.Length = "Sepal Length", Sepal.Width = "Sepal Width") %>%
  # Add a footer
  add_footer_lines("Iris dataset: Measurements of iris flowers from three species")
```

:::
