---
title: "Ex-2-1: Data Visualization"
format:
  html:
    toc: true
    toc-location: left
    number-sections: true
webr:
  packages: ['dplyr', 'ggplot2', 'datasets']
  cell-options:
    editor-font-scale: 0.7
    out-width: 100%
filters:
  - webr
---

# Basics

```{r}
#| include: false
library(dplyr)
library(ggplot2)
library(datasets)
```

## Exercise 1

**Objective:** Create a scatter plot of miles-per-gallon (`mpg`) against horsepower (`hp`), colored by the number of cylinders (`cyl`). Make the size of the points larger than the default (you can pick any value as long as it is larger than the default).

::: {.panel-tabset}

### Work here
```{webr-r}
data(mtcars)

#--- work here ---#



```

Here is the output you should be getting:

```{r}
#| echo: false
ggplot(data = mtcars) +
  geom_point(aes(x = hp, y = mpg, color = as.factor(cyl)), size = 2) 
```

### Answer
```{r, eval = FALSE}
#| code-fold: true
ggplot(data = mtcars) +
  geom_point(aes(x = hp, y = mpg, color = as.factor(cyl)), size = 2)
```

:::


## Exercise 2

**Objective:** Create a boxplot showing the distribution of petal lengths (`Petal.Length`) for each species (`Species`). Make the color of the borders of the boxes red.

::: {.panel-tabset}
### Work here
```{webr-r}
data(iris)

#--- work here ---#



```


Here is the output you should be getting:

```{r}
#| echo: false 
ggplot(data = iris) +
  geom_boxplot(aes(x = Species, y = Petal.Length, fill = Species), color = "red")
```


### Answer
```{r, eval = FALSE}
#| code-fold: true
ggplot(data = iris) +
  geom_boxplot(aes(x = Species, y = Petal.Length, fill = Species), color = "red") 
```

:::


## Exercise 3

**Objective:** Create a histogram of diamond prices (`price`) by `cut` (the fill color of the histogram differs by `cut`). Use `alpha = 0.6` so that the histograms are slightly transparent.

::: {.panel-tabset}
### Work here
```{webr-r}
data(diamonds)

#--- work here ---#



```


Here is the output you should be getting:

```{r}
#| echo: false
ggplot(data = diamonds) +
  geom_histogram(
    aes(x = price, fill = cut),
    alpha = 0.6,
    bins = 50
  )
```

### Answer
```{r, eval = FALSE}
#| code-fold: true
ggplot(data = diamonds) +
  geom_histogram(
    aes(x = price, fill = cut),
    alpha = 0.6,
    bins = 50
  )
```

:::

## Exercise 4

**Objective:** Create a line plot that shows the progression of weight (`weight`) over time (`Time`) for each of the chicks that were fed Diet 1 (Hint: you first need to filter the data so that you only have the observations that has `Diet == 1`). Make the line color dependent on `Chick`.

::: {.panel-tabset}
### Work here
```{webr-r}
data(ChickWeight)

#--- work here ---#



```

Here is the output you should be getting:

```{r}
#| echo: false
ggplot(data = ChickWeight %>% filter(Diet == 1)) +
  geom_line(aes(x = Time, y = weight, color = Chick))
```

### Answer
```{r, eval = FALSE}
#| code-fold: true
ggplot(data = ChickWeight %>% filter(Diet == 1)) +
  geom_line(aes(x = Time, y = weight, color = Chick))
```

:::


# Faceted figures

## Exercise 1

**Objective:** Create a scatter plot of miles-per-gallon (`mpg`) against weight (`wt`), colored by the number of gears (`gear`). Facet the plot by the number of cylinders (`cyl`).

::: {.panel-tabset}
### Work here
```{webr-r}
data(mtcars)

#--- work here ---#



```

Here is the output you should be getting:

```{r}
#| echo: false 
ggplot(data = mtcars) +
  geom_point(aes(x = wt, y = mpg, color = as.factor(gear)))  +
  facet_wrap(~cyl)
```

### Answer
```{r, eval = FALSE}
#| code-fold: true
ggplot(data = mtcars) +
  geom_point(aes(x = wt, y = mpg, color = as.factor(gear))) +
  facet_wrap(~cyl)
```

:::

## Exercise 2

**Objective:** Create a boxplot of diamond prices (`price`) for each diamond cut (`cut`). Facet the plot by diamond color (`color`).

::: {.panel-tabset}
### Work here
```{webr-r}
data(diamonds)

#--- work here ---#



```

Here is the output you should be getting:

```{r}
#| echo: false
ggplot(data = diamonds) +
  geom_boxplot(aes(x = cut, y = price, fill = cut)) +
  facet_wrap(~color)
```

### Answer
```{r, eval = FALSE}
#| code-fold: true
ggplot(data = diamonds) +
  geom_boxplot(aes(x = cut, y = price, fill = cut)) +
  facet_wrap(~color)
```

:::


## Exercise 3

**Objective:** Plot the progression of weight (`weight`) over time (`Time`) for each chick using a line plot. Facet the visualization by the diet type (`Diet`).

::: {.panel-tabset}
### Work here
```{webr-r}
data(ChickWeight)

#--- work here ---#



```

Here is the output you should be getting:

```{r}
#| echo: false 
ggplot(data = ChickWeight) +
  geom_line(aes(x = Time, y = weight, group = Chick)) +
  facet_wrap(~Diet) 
```

### Answer
```{r, eval = FALSE}
#| code-fold: true
ggplot(data = ChickWeight) +
  geom_line(aes(x = Time, y = weight, group = Chick)) +
  facet_wrap(~Diet)
```

:::

## Exercise 4

**Objective:** Create a scatter plot of highway miles-per-gallon (`hwy`) against engine displacement (`displ`). Facet the plot by the drive type (`drv`), with different panels for each type of drive (front-wheel drive, rear-wheel drive, and four-wheel drive).

::: {.panel-tabset}
### Work here
```{webr-r}
data(mpg)

#--- work here ---#



```

Here is the output you should be getting:

```{r}
#| echo: false 
ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy, color = class)) +
  facet_wrap(~drv)
```

### Answer
```{r, eval = FALSE}
#| code-fold: true
ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy, color = class)) +
  facet_wrap(~drv)
```

:::

## Exercise 5

**Objective:** Plot a boxplot showing the count of insects (`count`) for each spray type (`spray`). Facet the plot by spray type (`spary`).


::: {.panel-tabset}
### Work here
```{webr-r}
data(InsectSprays)

#--- work here ---#



```

Here is the output you should be getting: 

```{r}
#| echo: false 
ggplot(data = InsectSprays) +
  geom_boxplot(aes(x = spray, y = count, fill = spray)) +
  facet_wrap(~spray)
```

### Answer
```{r, eval = FALSE}
#| code-fold: true
ggplot(data = InsectSprays) +
  geom_boxplot(aes(x = spray, y = count, fill = spray)) +
  facet_wrap(~spray)
```

:::


# More exercises with emphasis on adjusting figure appearence

## Exercise 1

**Objective:** 
- Create a scatter plot of miles-per-gallon (`mpg`) against horsepower (`hp`).
- Use a color scheme that displays a gradient based on `hp`.
- Adjust the title text size and color.


::: {.panel-tabset}
### Work here
```{webr-r}
data(mtcars)

#--- work here ---#



```

Here is the output you should be getting:

```{r}
#| echo: false
ggplot(data = mtcars) +
  geom_point(aes(x = hp, y = mpg, color = hp)) +
  scale_color_gradient(low = "blue", high = "red") +
  theme(plot.title = element_text(size = 24, color = "darkgreen"))
```

### Answer
```{r, eval = FALSE}
#| code-fold: true
ggplot(data = mtcars) +
  geom_point(aes(x = hp, y = mpg, color = hp)) +
  scale_color_gradient(low = "blue", high = "red") +
  theme(plot.title = element_text(size = 24, color = "darkgreen"))
```

:::


## Exercise 2

**Objective:** 
- Generate a boxplot of sepal lengths (`Sepal.Length`) for each species (`Species`).
- Choose custom colors for the boxplot fills.
- Remove the legend entirely.

::: {.panel-tabset}
### Work here
```{webr-r}
data(iris)

#--- work here ---#



```

Here is the output you should be getting:

```{r}
#| echo: false 
ggplot(data = iris) +
  geom_boxplot(aes(x = Species, y = Sepal.Length, fill = Species)) +
  scale_fill_manual(values = c("Setosa" = "purple", "Versicolor" = "orange", "Virginica" = "green")) +
  theme(legend.position = "none")
```


### Answer
```{r, eval = FALSE}
#| code-fold: true
ggplot(data = iris) +
  geom_boxplot(aes(x = Species, y = Sepal.Length, fill = Species)) +
  scale_fill_manual(values = c("Setosa" = "purple", "Versicolor" = "orange", "Virginica" = "green")) +
  theme(legend.position = "none")
```

:::


## Exercise 3

**Objective:** 
- Design a histogram for diamond prices (`price`).
- Set bin colors to a specified hue, and the fill to another.
- Increase the x-axis and y-axis label text size.

::: {.panel-tabset}
### Work here
```{webr-r}
data(diamonds)

#--- work here ---#



```

Here is the output you should be getting:

```{r}
#| echo: false 
ggplot(data = diamonds) +
  geom_histogram(aes(x = price), color = "navy", fill = "lightpink", bins = 30) +
  theme(
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18)
  )
```

### Answer
```{r, eval = FALSE}
#| code-fold: true
ggplot(data = diamonds) +
  geom_histogram(aes(x = price), color = "navy", fill = "lightpink", bins = 30) +
  theme(
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18)
  )
```

:::

## Exercise 4

**Objective:** 
- Produce a scatter plot of highway miles-per-gallon (`hwy`) against engine displacement (`displ`).
- Modify the point shape based on drive type (`drv`).
- Rename the legend title and entries.


::: {.panel-tabset}
### Work here
```{webr-r}
data(mpg)
#--- work here ---#



```

Here is the output you should be getting:

```{r}
#| echo: false 
ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy, shape = drv, color = class), size = 3) +
  scale_shape_manual(
    name = "Drive Type",
    values = c("f" = 16, "r" = 17, "4" = 18),
    labels = c("Front", "Rear", "Four-wheel")
  )
```

### Answer
```{r, eval = FALSE}
#| code-fold: true
ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy, shape = drv, color = class), size = 3) +
  scale_shape_manual(
    name = "Drive Type",
    values = c("f" = 16, "r" = 17, "4" = 18),
    labels = c("Front", "Rear", "Four-wheel")
  )
```

:::


## Exercise 5

**Objective:** 
- Create a bar chart showing average tooth growth (`len`) by supplement type (`supp`).
- Change bar fill color based on dose (`dose`) using a bespoke palette.
- Position the legend at the bottom of the figure.


::: {.panel-tabset}
### Work here
```{webr-r}
data(ToothGrowth)

#--- work here ---#



```

Here is the output you should be getting:

```{r}
#| echo: false 
tooth_avg <- ToothGrowth %>%
  group_by(supp, dose) %>%
  summarize(average_len = mean(len))

ggplot(data = tooth_avg, ) +
  geom_bar(
    aes(x = supp, y = average_len, fill = as.factor(dose)),
    stat = "identity",
    position = "dodge"
  ) +
  scale_fill_manual(values = c("0.5" = "lightblue", "1.0" = "pink", "2.0" = "yellow")) +
  theme(legend.position = "bottom")
```

### Answer
```{r, eval = FALSE}
#| code-fold: true
tooth_avg <- ToothGrowth %>%
  group_by(supp, dose) %>%
  summarize(average_len = mean(len))

ggplot(data = tooth_avg, ) +
  geom_bar(
    aes(x = supp, y = average_len, fill = as.factor(dose)),
    stat = "identity",
    position = "dodge"
  ) +
  scale_fill_manual(values = c("0.5" = "lightblue", "1.0" = "pink", "2.0" = "yellow")) +
  theme(legend.position = "bottom")

```

:::